Columnar VCF for Biobank scale datasets using Zarr

### 2024-04-11 paper.tex
Update figs for new data

The data shown in Figure~\ref{fig-whole-matrix-compute} for Zarr and Savvy
is based on custom programs written using their respective APIs
to implement the \texttt{af-dist} operation. The Zarr program uses
the Zarr-Python package to iterate over the decoded chunks of the 
genotype matrix and classifies genotypes within a chunk using a 14 line Python
function, accelerated using the Numba JIT compliler~\cite{lam2015numba}.
The allele frequencies and genotype counts are then analysed to produce 
the final counts within the allele frequency bins with 9 lines of 
Python using NumPy~\cite{harris2020array} functions. Remarkably, this 
short and simple Python program is substantially faster than the 
equivalent compiled C using \texttt{htslib} APIs on BCF (6 hours
vs 20 hours for 1 million samples). The fastest method is the 
C++ program written using the Savvy API. This would largely seem
to be due to Savvy's excellent genotype decoding performance
(up to 5.7GiB/s vs 1.2GiB/s for Zarr on this dataset;
Figure~\ref{fig-whole-matrix-decode}). Different choices 
of compressor and filters (zstandard with bit shuffle, in this case) 
would likely lead to improved decoding performance in Zarr,
at the expense of somewhat larger files.

>>>>

The data shown in Figure~\ref{fig-whole-matrix-compute} for Zarr and Savvy
is based on custom programs written using their respective APIs
to implement the \texttt{af-dist} operation. The Zarr program uses
the Zarr-Python package to iterate over the decoded chunks of the 
genotype matrix and classifies genotypes within a chunk using a 14 line Python
function, accelerated using the Numba JIT compliler~\cite{lam2015numba}.
The allele frequencies and genotype counts are then analysed to produce 
the final counts within the allele frequency bins with 9 lines of 
Python using NumPy~\cite{harris2020array} functions. Remarkably, this 
short and simple Python program is substantially faster than the 
equivalent compiled C using \texttt{htslib} APIs on BCF (6 hours
vs 20 hours for 1 million samples). The fastest method is the 
C++ program written using the Savvy API. This would largely seem
to be due to Savvy's excellent genotype decoding performance
(up to 6.5GiB/s vs 1.2GiB/s for Zarr on this dataset;
Figure~\ref{fig-whole-matrix-decode}). Different choices 
of compressor and filters (zstandard with bit shuffle, in this case) 
would likely lead to improved decoding performance in Zarr,
at the expense of somewhat larger files.
